---
title: "ST558 Project1"
author: "Owen Snyder"
date: '2022-06-16'
output: 
  html_document:
    toc: true
---

Create render function
```{r eval=FALSE}
library(rmarkdown)
render(
 output_format = "github_document",
  output_dir = "_posts",
  output_options = list(
  html_preview = FALSE
)
)
 

 output_format = "github_document"
  output_dir = "_posts"
  output_options = list(
  html_preview = FALSE
  )
```



# Necessary Packages
```{r}
library(tidyverse)
library(ggplot2)
library(httr)
library(jsonlite)

```

dont need?, i.e. dont need to include, echo = F
```{r}
fin.data <- GET("https://api.polygon.io/v2/aggs/ticker/AAPL/range/1/day/2020-06-01/2020-06-17?apiKey=O5bhHpJ7TFFBgq7bUJp8TA7Yi7uDLMIe")

str(fin.data, max.level = 1)

parsed <- fromJSON(rawToChar(fin.data$content))
str(parsed, max.level = 1)
```


```{r}
rawToChar(fin.data$content)
fromJSON(readLines("https://api.polygon.io/v1/open-close/AAPL/2020-10-14?adjusted=true&apiKey=O5bhHpJ7TFFBgq7bUJp8TA7Yi7uDLMIe"))
```

# Function for Daily Open/Close for Stocks
Create function to grab Daily Open/Close for stocks
```{r}
getDailyOC <- function(stocksTicker,date,adjusted){
  adj <- ifelse(adjusted, "true", "false")
  link <- paste0("https://api.polygon.io/v1/open-close/",stocksTicker,"/",date,"?adjusted=",adj,"&apiKey=O5bhHpJ7TFFBgq7bUJp8TA7Yi7uDLMIe")
  
  fromJSON(readLines(link,warn = FALSE))
 
}
## test
getDailyOC(stocksTicker = "SPY", date = "2021-10-12", adjusted = TRUE)

```

grab Crypto data (open/close)
# Function for Daily Open/Close for Crypto
```{r}
## use this function if you want to know the open/close prices for a given crypto currecy on a singular day.
## NOTE: do this for all^

getCryptoOC <- function(symbol, to, date, adjusted,...){
  adj <- ifelse(adjusted, "true", "false")
  link <-  paste0("https://api.polygon.io/v1/open-close/crypto/",symbol,"/",to,"/",date,"?adjusted=",adj,"&apiKey=O5bhHpJ7TFFBgq7bUJp8TA7Yi7uDLMIe")
  
  fromJSON(readLines(link,warn = FALSE))
}

getCryptoOC(symbol = "ETH", to = "USD", date = "2021-10-12", adjusted = TRUE)
```


grab Aggregate (stocks)
# Function for Aggregates(Bars) Stock Data
```{r}

getAgg <- function(stocksTicker,multiplier,timespan="day",from,to,adjusted,sort="desc",limit=5000){
  
  adj <- ifelse(adjusted, "true", "false")
  multiplier <- as.integer(multiplier)
  timespan <- match.arg(timespan,choices = c("minute", "hour", "day", "week", "month", "quarter",                                                   "year"))
  sort <- match.arg(sort, choices = c("asc", "desc"))
  limit <- as.integer(limit)
  if (limit > 50000)
    stop("limit must be less than 50000")
  if(limit < 1)
    stop("limit must be gt or eq 1")
  
    
  link <-  paste0("https://api.polygon.io/v2/aggs/ticker/",stocksTicker,"/range/",multiplier,"/"
                  ,timespan,"/",from,"/",to,"?adjusted=",adj,"&sort=",sort,"&limit=",limit,"&apiKey=O5bhHpJ7TFFBgq7bUJp8TA7Yi7uDLMIe")
  
  output <- fromJSON(readLines(link,warn = FALSE))
  
  return(output$results)
  
  ## this is only printing out the tibble of data you request
}

#get.appl <- getAgg("SPY", multiplier = 1, timespan = "day", from = "2020-10-01", to= "2022-06-15",adjusted = TRUE, sort = 'desc', limit = 6000)


## Now i am renaming column names 
colnames(get.appl) <- c('Volume', 'WeightedVolume', "Open", "Close", "HighestPrice",
                        "LowestPrice", "UnixTimes(msec)", "NumTransactions")

get.appl
```


# Function for Aggregates(Bars) Crypto Data
```{r}
## note: the pre-inputted arguments are for common defualts. they can be changed per
## the users request
getAggCryp <- function(cryptoTicker,multiplier,timespan="day",from,to,adjusted,
                       sort="desc", limit=5000){
  
  adj <- ifelse(adjusted, "true", "false")
  multiplier <- as.integer(multiplier)
  timespan <- match.arg(timespan,choices = c("minute", "hour", "day", "week", "month", "quarter",                                                   "year"))
  sort <- match.arg(sort, choices = c("asc", "desc"))
  limit <- as.integer(limit)
  if (limit > 50000)
    stop("limit must be less than 50000")
  if(limit < 1)
    stop("limit must be gt or eq 1")
  
  link <- paste0("https://api.polygon.io/v2/aggs/ticker/",cryptoTicker,"/range/",multiplier,"/",timespan,"/",from,"/",to,"?adjusted=",adj,"&sort",sort,"&limit=",limit,"&apiKey=O5bhHpJ7TFFBgq7bUJp8TA7Yi7uDLMIe")
  
  output <- fromJSON(readLines(link,warn = FALSE))
  return(output$results)
  
}
## grab Ethereum
get.eth <- getAgg("X:ETHUSD", multiplier = 1, timespan = "day", from = "2021-10-01", to= "2022-12-31",adjusted = TRUE, sort = 'desc', limit = 5000)
## change col names
get.eth <- getAggCryp("X:ETHUSD", multiplier = 1, timespan = "day", from = "2021-10-01", to= "2022-12-31",adjusted = TRUE, sort = 'desc', limit = 5000)

```


function to see all types of asset based on location (US or global)
# Function to find all Exchanges  the API carries
```{r}
getEXCH <- function(asset_class,locale){
  
  asset_class <- match.arg(asset_class, choices = c("stocks","options","crypto","fx"))
  locale <- match.arg(locale, choices = c("us","global"))
  
  link <- paste0("https://api.polygon.io/v3/reference/exchanges?asset_class=",asset_class,"&apiKey=O5bhHpJ7TFFBgq7bUJp8TA7Yi7uDLMIe")
  
  output <- fromJSON(readLines(link,warn = FALSE))
}
## hmmm....
get.all <- getEXCH(asset_class = "stocks", locale = "us")


```

Last function...
# Function to....
```{r}
##hmmm...
## i dont think i need another...wait yeah i do 
```


LETS COMPARE CRYTO AND STOCK PRICES FOR A GIVEN TIME PERIOD

“Bars” refer to a data representation that contains the most basic information about price movements of a financial asset.

NEW VARIABLE SHOULD BE  RETURNS and LOG RETURNS

# Exploratory Data Analysis
Modeling stock and crypto performance can be a challenging task as there are many different measures we can assess. You may want to analyze one stock or a portfolio of stocks. and You may want to look at how risky these positions are.  

In this EDA, I will pull data from the S&P 500 and different cryptocurrencies such as Bitcoin, Etherieum and Solana to see how they perform against my chosen date window. I want to analyze if there is any similarity between stock market trends and crypto trends 

NOTE: It is important to recogtnize the major market differences between the tradional stock market and crypto currencies. Crypto is a 24/7 market

First, I will extract the necessary data i am going to work with via my created functions from above.

## Extract SPY Data

```{r}
## use the getAgg function to pull S&P 500 data 

data.spy <- getAgg("SPY", multiplier = 1, timespan = "day", from = "2020-10-01", to= "2022-06-15",adjusted = FALSE, sort = 'asc', limit = 5000)
## Now i am renaming column names 
colnames(data.spy) <- c('Volume', 'WeightedVolume', "Open", "Close", "HighestPrice",
                        "LowestPrice", "UnixTimes(msec)", "NumTransactions")
data.spy

##plot(data.spy$Close)

## create a new variable for Log Returns 
data.spy %>% mutate(LogRet = log(Close) - log(lag(Close))) %>% replace(is.na(.), 0)
SPY <- data.spy %>% mutate(LogRet = log(Close) - log(lag(Close))) %>% replace(is.na(.), 0) %>% 
       mutate(Symbol = 'SPY')

## NotE: set na = 0 because there are no returns on the first day/observation

#data.spy.spy <- data.spy %>% as_tibble() %>% mutate(LogRet = log(Close) - log(lag(Close))) %>% replace(is.na(.), 0)

#plot(data.spy.spy$LogRet)
```

## Extract Crypto Data

```{r}
## use the getAggCryp function to pull Bitcoin Data

data.btc <- getAggCryp("X:BTCUSD", multiplier = 1, timespan = "day", from = "2020-10-01", to= "2022-06-15",                           adjusted = FALSE, sort = 'asc', limit = 5000)
## Now change column names again
colnames(data.btc) <- c('Volume', 'WeightedVolume', "Open", "Close", "HighestPrice",
                        "LowestPrice", "UnixTimes(msec)", "NumTransactions")
data.btc

plot(data.btc$Close)

## can graph num transactions vs close price

data.eth <- getAggCryp("X:ETHUSD", multiplier = 1, timespan = "day", from = "2020-10-01", to= "2022-06-15",                           adjusted = FALSE, sort = 'asc', limit = 5000)
plot(data.eth$c)
```




## Numerical Summaries 

I first want to display some simple summary statistics for the stocks and the cryptocurrecnies.

```{r}
## Summary for SPY data 
SPY %>% summarise(meanVol = mean(Volume), meanOpen = mean(Open), meanClose = mean(Close))

```
