---
title: "ST558 Project1"
author: "Owen Snyder"
date: '2022-06-16'
output: 
  html_document:
    toc: true
---

Create render function
```{r eval=FALSE}

rmarkdown::render('README.md',
 output_format = "github_document",
 #output_dir = "_posts",
  output_options = list(
  html_preview = TRUE
)
)
 

 output_format = "github_document"
  output_dir = "_posts"
  output_options = list(
  html_preview = FALSE
  )
```



# Necessary Packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(httr)
library(jsonlite)
library(lubridate)
library(bizdays)

```

dont need?, i.e. dont need to include, echo = F
```{r}
fin.data <- GET("https://api.polygon.io/v2/aggs/ticker/AAPL/range/1/day/2020-06-01/2020-06-17?apiKey=O5bhHpJ7TFFBgq7bUJp8TA7Yi7uDLMIe")

str(fin.data, max.level = 1)

parsed <- fromJSON(rawToChar(fin.data$content))
str(parsed, max.level = 1)
```


```{r warning=FALSE, echo=FALSE}
rawToChar(fin.data$content)
fromJSON(readLines("https://api.polygon.io/v1/open-close/AAPL/2020-10-14?adjusted=true&apiKey=O5bhHpJ7TFFBgq7bUJp8TA7Yi7uDLMIe"))
```

# Function for Daily Open/Close for Stocks

```{r}
## This function will take in values of the ticker of a given stock, the date (day) you want to 
## look at, and whether or not you want adjusted splits

getDailyOC <- function(stocksTicker,date,adjusted){
  adj <- ifelse(adjusted, "true", "false") ##setting adjusted to be only true or false
  link <- paste0("https://api.polygon.io/v1/open-close/",stocksTicker,"/",date,"?adjusted=",adj,"&apiKey=O5bhHpJ7TFFBgq7bUJp8TA7Yi7uDLMIe")
  ## the link to connect to the API endpoint
  fromJSON(readLines(link,warn = FALSE))
 ## here we are converting a JSON object into a usable R object and suppressing warnings
}
## test the function with a ticker
## getDailyOC(stocksTicker = "SPY", date = "2021-10-12", adjusted = TRUE)

```


# Function for Daily Open/Close for Crypto

```{r}
## This function will take in values for the symbol of a given cryptocurrency, the currency you want 
## the crypto to be converted into, the date (day) you want to look at, and  and whether or not you want 
## adjusted splits

getCryptoOC <- function(symbol, to, date, adjusted,...){
  adj <- ifelse(adjusted, "true", "false") ##setting adjusted to be only true or false
  link <-  paste0("https://api.polygon.io/v1/open-close/crypto/",symbol,"/",to,"/",date,"?adjusted=",adj,"&apiK                    ey=O5bhHpJ7TFFBgq7bUJp8TA7Yi7uDLMIe")
  ## the link to connect to the API endpoint
  fromJSON(readLines(link,warn = FALSE))
  ## here we are converting a JSON object into a usable R object and suppressing warnings
}

## test the function with a crypto ticker
##getCryptoOC(symbol = "ETH", to = "USD", date = "2021-10-12", adjusted = TRUE)
```


# Function for Aggregates(Bars) Stock Data
```{r}
## this function will take in a stocks ticker, multiplier, time span, date range values (from and to),
## adjusted splits, a sort option, and a limit value
## Note: the pre-inputted arguments are for common defaults. they can be changed per the users request

getAgg <- function(stocksTicker,multiplier,timespan="day",from,to,adjusted,sort="desc",limit=5000){
  
  adj <- ifelse(adjusted, "true", "false") ## setting adjusted to be only true or false
  multiplier <- as.integer(multiplier) ## multiplier must be an integer 
  timespan <- match.arg(timespan,choices = c("minute", "hour", "day", "week", "month", "quarter", "year"))
  ## provide a list of time span choices the URL and function will be able to include
  sort <- match.arg(sort, choices = c("asc", "desc"))
  ## establish the sort of how you want data to be printed 
  limit <- as.integer(limit)
  if (limit > 50000)
    stop("limit must be less than 50000")
  if(limit < 1)
    stop("limit must be greater than or equal to 1")
  ## set limit conditions and stop the function if they fail
    
  link <-  paste0("https://api.polygon.io/v2/aggs/ticker/",stocksTicker,"/range/",multiplier,"/"
                  ,timespan,"/",from,"/",to,"?adjusted=",adj,"&sort=",sort,"&limit=",limit,"&apiKey=O5bhHpJ7TFF                   Bgq7bUJp8TA7Yi7uDLMIe")
  ## the link to connect to the API endpoint
  output <- fromJSON(readLines(link,warn = FALSE))
  ## here we are converting a JSON object into a usable R object and suppressing warnings
  
  return(output$results)
  ## return only the results 
  
}

## test the function
## get.appl <- getAgg("SPY", multiplier = 1, timespan = "day", from = "2020-10-01", to= "2022-06-15",adjusted = TRUE, sort = 'desc', limit = 6000)

```


# Function for Aggregates(Bars) Crypto Data

```{r}
## this function will take in a crypto ticker, a multiplier value, a time span, date ranges (from and to)
## adjusted splits, a sort option, and limit value
getAggCryp <- function(cryptoTicker,multiplier,timespan="day",from,to,adjusted,
                       sort="desc", limit=5000){
  
  adj <- ifelse(adjusted, "true", "false") ## setting adjusted to be only true or false
  multiplier <- as.integer(multiplier) ## multiplier must be an integer
  timespan <- match.arg(timespan,choices = c("minute", "hour", "day", "week", "month", "quarter", "year"))
  ## provide a list of time span choices the URL and function will be able to include
  sort <- match.arg(sort, choices = c("asc", "desc"))
  ## establish the sort of how you want data to be printed
  limit <- as.integer(limit)
  if (limit > 50000)
    stop("limit must be less than 50000")
  if(limit < 1)
    stop("limit must be greater than or equal to 1")
  
  
  link <- paste0("https://api.polygon.io/v2/aggs/ticker/",cryptoTicker,"/range/",multiplier,"/"
                 ,timespan,"/",from,"/",to,"?adjusted=",adj,"&sort",sort,"&limit=",limit,"&apiKey=O5bhHpJ7TFFBg                  q7bUJp8TA7Yi7uDLMIe")
  ## the link to connect to the API endpoint
  output <- fromJSON(readLines(link,warn = FALSE))
  return(output$results) ## return only the results 
}

## test function
## get.eth <- getAgg("X:ETHUSD", multiplier = 1, timespan = "day", from = "2021-10-01", to= "2022-12-31",
##                  adjusted = TRUE, sort = 'desc', limit = 5000)
```


# Function to find all Exchanges the API carries

```{r}
## this function will take in an asset class (options specified below) and a locale (location)

getEXCH <- function(asset_class,locale){
  
  asset_class <- match.arg(asset_class, choices = c("stocks","options","crypto","fx"))
  ## choices for asset_class must be these four
  locale <- match.arg(locale, choices = c("us","global"))
  ## locale must be US or global
  link <- paste0("https://api.polygon.io/v3/reference/exchanges?asset_class=",asset_class,
                 "&apiKey=O5bhHpJ7TFFBgq7bUJp8TA7Yi7uDLMIe")
  ## the link to connect to the API endpoint
  output <- fromJSON(readLines(link,warn = FALSE))
  return(output$results) ## return only the results section
  
}

## test the function
## get.all <- getEXCH(asset_class = "stocks", locale = "us")


```


# Function to find Grouped Daily(Bars) for All Stocks

```{r}
## Get the daily open, high, low, and close (OHLC) for the entire stocks/equities markets

getGroupBars <- function(date, adjusted, include_otc){
  adj <- ifelse(adjusted, "true", "false") ## setting adjusted to be only true or false
  otc <- ifelse(include_otc, "true", "false") ## setting include_otc to be only true or false
  link <- paste0("https://api.polygon.io/v2/aggs/grouped/locale/us/market/stocks/",date,"?adjusted=",adj,
                 "&apiKey=O5bhHpJ7TFFBgq7bUJp8TA7Yi7uDLMIe")
  ## the link to connect to the API endpoint
  output <- fromJSON(readLines(link,warn = FALSE))
  return(output$results) 
  ## return only the results
}

## test function
## getGroupBars(date = "2021-10-12", adjusted = TRUE, include_otc = TRUE)

```


LETS COMPARE CRYTO AND STOCK PRICES FOR A GIVEN TIME PERIOD

“Bars” refer to a data representation that contains the most basic information about price movements of a financial asset.

NEW VARIABLE SHOULD BE  RETURNS and LOG RETURNS

# Exploratory Data Analysis
Modeling stock and crypto performance can be a challenging task as there are many different measures we can assess. You may want to analyze one stock or a portfolio of stocks. and You may want to look at how risky these positions are.  

In this EDA, I will pull data from the S&P 500 and different cryptocurrencies such as Bitcoin, Etherieum and Solana to see how they perform against my chosen date window. I want to analyze if there is any similarity between stock market trends and crypto trends 

NOTE: It is important to recogtnize the major market differences between the tradional stock market and crypto currencies. Crypto is a 24/7 market

First, I will extract the necessary data i am going to work with via my created functions from above.

## Extract S&P 500 Data (SPY)

```{r}
## use the getAgg function to pull S&P 500 data 

data.spy <- getAgg("SPY", multiplier = 1, timespan = "day", from = "2021-06-15", to= "2022-06-15",adjusted = FALSE, sort = 'asc', limit = 5000)
## Now i am renaming column names 
colnames(data.spy) <- c('Volume', 'WeightedVolume', "Open", "Close", "HighestPrice",
                        "LowestPrice", "UnixTimes(msec)", "NumTransactions")
data.spy

##plot(data.spy$Close)

## create a new variable for Log Returns 
data.spy %>% mutate(LogRet = log(Close) - log(lag(Close))) %>% replace(is.na(.), 0)
SPY <- data.spy %>% mutate(LogRet = log(Close) - log(lag(Close))) %>% replace(is.na(.), 0) %>% 
       mutate(Symbol = 'SPY') %>% mutate(Date = range)

## NotE: set na = 0 because there are no returns on the first day/observation

#data.spy.spy <- data.spy %>% as_tibble() %>% mutate(LogRet = log(Close) - log(lag(Close))) %>% replace(is.na(.), 0)

#plot(data.spy.spy$LogRet)
```

## Extract Vanguard 500 Index Fund (VOO)

```{r}
data.voo <- getAgg("VOO", multiplier = 1, timespan = "day", from = "2021-06-15", to= "2022-06-15",adjusted = FALSE, sort = 'asc', limit = 5000)
## Now i am renaming column names 
colnames(data.voo) <- c('Volume', 'WeightedVolume', "Open", "Close", "HighestPrice",
                        "LowestPrice", "UnixTimes(msec)", "NumTransactions")

data.voo

## create a new variable for Log Returns 
VOO <- data.voo %>% mutate(LogRet = log(Close) - log(lag(Close))) %>% replace(is.na(.), 0) %>% 
       mutate(Symbol = "VOO")

#VOO %>% select(NumTransactions)
#summary(VOO$NumTransactions)
## maybe find which ones are fewer and most
```

## Extract REIT data

```{r}
## Camden Property Trust
data.cpt <- getAgg("CPT", multiplier = 1, timespan = "day", from = "2021-06-15", to= "2022-06-15",adjusted = FALSE, sort = 'asc', limit = 5000)
## Now i am renaming column names 
colnames(data.cpt) <- c('Volume', 'WeightedVolume', "Open", "Close", "HighestPrice",
                        "LowestPrice", "UnixTimes(msec)", "NumTransactions")
data.cpt

## add everything else with mutate 
```



Combine stock data
```{r}
all.stocks <- bind_rows(SPY,VOO)
length(SPY)

```

Date formatting 
```{r}
start.date <- ymd("2021-06-15")
end.date <- ymd("2022-06-15")
range <- seq(start.date, end.date,"days")
head(range, 40)
fin.quarters <- quarters(range)

omg <- range[lubridate::wday(range) %in% 2:6] # remove weekdays 
length(range[lubridate::wday(range) %in% 2:6])

work.days <- !is.holiday(omg)

final.stock.date <- omg[-c("2022-05-30")]

```


## Extract Crypto Data

```{r}
## use the getAggCryp function to pull Bitcoin Data

data.btc <- getAggCryp("X:BTCUSD", multiplier = 1, timespan = "day", from = "2021-06-15", to= "2022-06-15",                           adjusted = FALSE, sort = 'asc', limit = 5000)
## Now change column names again
colnames(data.btc) <- c('Volume', 'WeightedVolume', "Open", "Close", "HighestPrice",
                        "LowestPrice", "UnixTimes(msec)", "NumTransactions")

BTC <- data.btc %>% mutate(LogRet = log(Close) - log(lag(Close))) %>% replace(is.na(.), 0) %>% 
       mutate(Symbol = "BTC") %>% mutate(Date = range, Quarters = fin.quarters)


## can graph num transactions vs close price


data.eth <- getAggCryp("X:ETHUSD", multiplier = 1, timespan = "day", from = "2021-06-15", to= "2022-06-15",                           adjusted = FALSE, sort = 'asc', limit = 5000)
colnames(data.eth) <- c('Volume', 'WeightedVolume', "Open", "Close", "HighestPrice",
                        "LowestPrice", "UnixTimes(msec)", "NumTransactions")

ETH <- data.eth %>% mutate(LogRet = log(Close) - log(lag(Close))) %>% replace(is.na(.), 0) %>% 
       mutate(Symbol = "ETH") %>% mutate(Date = range, Quarters = fin.quarters) %>% 
       mutate(TransCategory = ifelse(NumTransactions %in% 800000:1852104, "Many",
                              ifelse(NumTransactions %in% 500000:800000, "Average",
                              ifelse(NumTransactions %in% 42510:500000, "Few", "NONE"))))
```

Combine Crypto
```{r}
all.crypto <- bind_rows(BTC,ETH)
all.c <- merge(BTC,ETH, by = 'Date', all.x = TRUE)
all.poo <- merge(BTC,ETH, by.x = 'Date', 'Quarter', by.y = 'Date', 'Quarter', all.x = FALSE, all.y = FALSE )
## table but need counts 

#left_join(BTC,ETH, by = "Date","Quarters")
summary(ETH$NumTransactions)

ETH %>% mutate(TransCategory = ifelse(NumTransactions %in% 750000:1852104, "Many",
                                      ifelse(NumTransactions %in% 450000:750000, "Average",
                                      ifelse(NumTransactions %in% 42510:450000, "Few", "NONE"))))
```

contingency tables
```{r}
hist(ETH$NumTransactions)
table(ETH$TransCategory, ETH$Quarters)
```


## Numerical Summaries 

I first want to display some simple summary statistics for the stocks and the cryptocurrecnies.

```{r}
## Summary for SPY data 
SPY %>% summarise(meanVol = mean(Volume), meanOpen = mean(Open), meanClose = mean(Close))

```



Graphs:

Histograms
```{r}

p <- ggplot(data = BTC, aes(x = BTC$LogRet))
p+geom_histogram() + geom_density()

p2 <- ggplot(data = ETH, aes(x = LogRet))
p2+geom_histogram() + geom_density()

p3 <- ggplot(data = SPY, aes(x = LogRet))
p3+geom_histogram()

p4 <- ggplot(data = VOO, aes(x = LogRet))
p4+geom_histogram()
```


Boxplots:
```{r}
#bp1 <- ggplot(data = all.c, x= all.c$NumTransactions.y, y = all.c$Open.y )
#bp1 + geom_boxplot()

## boxplot for quarters and open prices
bp1 <- ggplot(BTC, aes(x = Quarters, y = Close))
bp1 + geom_boxplot()


bp2 <- ggplot(ETH, aes(x = Quarters, y = Close))
bp2 + geom_boxplot()
```



Time Series Graph for Bitcoin
```{r}
ggplot(data = BTC, aes(x=Date)) + 
  geom_line(aes(y=LogRet)) + 
  labs(title="Time Series Plot on Log Returns for Bitcoin", 
       subtitle="Returns Based on Close Price", 
       y="Log Returns") +  # title and caption
  theme(axis.text.x = element_text(angle = 90, vjust=0.5),  # rotate x axis text
        panel.grid.minor = element_blank())  # turn off minor grid 

```

Time Series Graph for Ethereum
```{r}
ggplot(data = ETH, aes(x=Date)) + 
  geom_line(aes(y=LogRet)) + 
  labs(title="Time Series Plot on Log Returns for Ethereum", 
       subtitle="Returns Based on Close Price", 
       y="Log Returns") +  
  theme(axis.text.x = element_text(angle = 90, vjust=0.5))  
        #panel.grid.minor = element_blank())  # turn off minor grid 

## CHECK GRIDS
```

Here, we notice that ETH is much nosier than Bitcoin.




```{r}

plot(ETH$Open,ETH$NumTransactions)
plot(ETH$NumTransactions, ETH$Open)
plot(ETH$Volume, ETH$Open)


```
